Rising from the Ashes: How the Milky Way Got Its Scars

### 2024-04-21 paper/ms.tex
Update abstract

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
  Stars are thought to inherit the elemental abundances of their birth cloud and, for the most part, to not change their surface abundances over time. Stellar surface abdundances are, therefore, a promising avenue to probe the time evolution of the abundance distribution of the Galaxy's gas phase. A particularly useful compression of the high-dimensional abundance plane is to consider the distribution of $\alpha$ elements for a given total metallicity, usually tracked by iron. It is well-known that this distribution of $\alpha$ is bimodal, for which there is a growing consensus that the low-$\alpha$ sequence diverges from the high-$\alpha$ sequence as a result of a merger. The general idea is that fresh, pristine gas from the merger causes a sudden change in the abundance structure of the disk, such that the gas spends relatively little time in the valley of bimodality. However, we argue that it is also possible that when the Galaxy's gas phase was in the valley that it was quiescent. This relaxes the necessity for a rapid abundance transition. A preceding starburst, which may cause the quiescent phase, would also be a culprit in the abundance transition.

  In this work, we propose the phoenix hypothesis: the bimodality is caused by a brief ($\sim300\,\Myr$) period of quiescent star formation, possibly preceded by a starburst. The bimodality is a natural consequence of stellar populations enriching the Galactic disk during a period in which no stars are forming. Our proposal is consistent with a merger, but it is agnostic to the exact mechanism by which the quiescent period arises. Our proposal predicts no overlap in age between the high- and low-$\alpha$ sequence at a given birth radius.
\end{abstract}

>>>>

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
  The nuclear abundance distribution of stars encodes the history of the gas-phase abundance in the Milky Way. Without a large, unbiased sample of stellar ages, the exact timing and nature of this history must be \textit{inferred} from the abundances. In the two-dimensional plane of \alphaFe-\FeH, it is now clear that two separate populations exist - the low-$\alpha$ and high-$\alpha$ sequences. A restatement of this fact is that the distribution of \alphaFe{} is bimodal at a fixed Fe. Structure in the nuclear abundance distribution can arise from many processes - proposals include radial migration, high-redshift clump formation, and various effects associated with galaxy mergers, among others. In this work, we demonstrate another possible avenue for structure formation with clear observational predictions. We propose that the Galaxy underwent a starburst followed by a brief (hundreds of Myr) quiescent phase. A natural consequence of this quiescent phase is that stars in the valley of the bimodality do not form because: (1) the absence of enrichment from high-mass stars leads to a rapid reduction in \alphaFe{}, and (2) any time the gas spends in the abundance valley is deemphasized in the present day distribution because the star formation rate is lower. With a set of idealized merger simulations, we demonstrate the feasibility of this proposal. The starburst and quiescent phase can result from a merger, but it is not a necessary component of the theory. This ``phoenix hypothesis'' predicts a gap in stellar ages at a fixed \FeH{} and that stars which form directly after this gap would have lower \alphaFe{} than stars which form slightly later.
\end{abstract}
